## Loading and preprocessing the data
```{r}
fitdata <- read.csv("activity.csv", header = TRUE, sep = ",", na.strings = "NA")
```
Explore the data
```{r}
summary(fitdata)
str(fitdata)
head(fitdata)
```
Convert the *date* variable to a Date class and the *interval* variable to a factor:
```{r}
fitdata$date <- as.Date(fitdata$date, format = "%Y-%m-%d")
fitdata$interval <- factor(fitdata$interval)
```

Subsetting the dataset to ignore missing values
```{r}
NA_index <- is.na(as.character(fitdata$steps))
noNA <- fitdata[!NA_index,]
head(noNA)
```
Aggregating the number of steps taken each day
Creating a data frame for steps taken each day
```{r}
stepsEachDay <- aggregate(steps ~date, data = noNA, sum)
```
Adding column names for created data frame
```{r}
colnames(stepsEachDay) <- c("date", "steps")
```
Histogram for the total number of steps taken each day
```{r}
hist(as.numeric(stepsEachDay$steps), breaks = 20, col = "red", 
     xlab = "Number of Steps", main= "Total Number of Steps Taken Each Day")
```
Copy plot
```{r}
dev.copy(png,"plot1.png", width=480, height=480)
```
Sign out
```{r}
dev.off()
```
## Mean of steps each day
```{r}
mean(stepsEachDay$steps)
```
Median of steps each day
```{r}
median(stepsEachDay$steps)
```
Calculating the average
```{r}
steps_per_interval <- aggregate(noNA$steps, by=list(interval=noNA$interval), 
                                FUN=mean)
```
Adding the columns names
```{r}
colnames(steps_per_interval) <- c("interval", "average_steps")
```
## The average daily activity pattern 
```{r}
plot(as.integer(levels(steps_per_interval$interval)), 
     steps_per_interval$average_steps, type="l",
     xlab = "Interval", ylab = "Average Number of Steps", 
     main = "Average Daily Activity Pattern",  col ="purple")
```
The maximum number of average steps
```{r}
maxSteps <- max(steps_per_interval$average_steps)
maxSteps
```
The 5-min interval containing the max number of steps
```{r}
interval_max_steps<-steps_per_interval[which.max(steps_per_interval$average_steps),]$interval
interval_max_steps
```
## Imputing Missing Values
For the steps variable:
```{r}
sum(is.na(as.character(fitdata$steps)))
```
For the date variable:
```{r}
sum(is.na(as.character(fitdata$date)))
```
For the interval variable:
```{r}
sum(is.na(as.character(fitdata$interval)))
```
Total number of missing values = 2304  

Replacing missing values by the mean of that 5-minute interval.  
Finding the indices of missing values (NAs)
```{r}
NA_index <- which(is.na(as.character(fitdata$steps)))
complete_data <- fitdata
```
Imputing missing values using the mean for that 5-minute interval
```{r}
complete_data[NA_index,]$steps <- unlist(lapply(NA_index, FUN=function(NA_index){
       steps_per_interval[fitdata[NA_index,]$interval==
                                 steps_per_interval$interval,]$average_steps
}))
```
Verifying complete data with the summary and str methods
```{r}
summary(complete_data)
str(complete_data)
```
Creating a data frame with the steps taken for each day
```{r}
steps_each_day_complete <- aggregate(steps ~ date, data = complete_data, sum)
```
Adding column names to the created data frame
```{r}
colnames(steps_each_day_complete) <- c("date", "steps")
```
Making the histogram
```{r}
hist(as.numeric(steps_each_day_complete$steps), breaks = 20, col = "orange", 
     xlab = "Number of Steps", main= "Total Number of Steps Taken Each Day")
```
Copy plot
```{r}
dev.copy(png,"plot2.png", width=480, height=480)
```
Sign out
```{r}
dev.off()
```




